---
title: "Capstone.Cleaning"
output: pdf_document
---
```{r}
library(rgdal)
library(raster)
library(ggmap)
library(lubridate)
library(sf)
library(ggplot2)
library(maps)
library(mapdata)
library(sp)
```


```{r}
#Read in cameras dataset
cameras<-read.csv("/Users/BeauSwallow/Downloads/Traffic_Camera.csv")
#Rename X & Y variables to Longitude and Latitude
cameras$Longitude<-cameras$X
cameras$Latitude<-cameras$Y
head(cameras)
```

```{r}
crashes<-read.csv("/Users/BeauSwallow/Downloads/Crashes_in_DC.csv")
#Clean up date column
crashes$REPORTDATE<-as.Date(crashes$REPORTDATE)
crashes$FROMDATE<-as.Date(crashes$FROMDATE)
#Subset by date for 2016-2019 data
crashes <- subset(crashes, REPORTDATE > as.Date("2016-01-01") )
head(crashes)
```

#Read in DC shape file with ward boundaries
```{r}
wardmap <- st_read(
  "/Users/BeauSwallow/Downloads/Ward_from_2012/Ward_from_2012.shp")
```
#Plot DC shapefile with ward boundaries
```{r}
ggplot() + 
  geom_sf(data = wardmap, size = 3, color = "black", fill = "cyan1") + 
  
  ggtitle("wardmap") + 
  coord_sf()
```
#Plotting cameras over it
```{r}
ggplot() + 
  geom_sf(data = wardmap, size = 3, color = "black", fill = "cyan1") + 
  coord_sf()+
geom_point(aes(x = Longitude, y = Latitude), data = cameras,
 alpha = .5, color="darkred", size = 0.5) +
ggtitle("Traffic Cameras in DC")
```
#DC Map using ggmap library
```{r}
myLocation <- c(-77.1198, 38.79164, -76.90915, 38.99597)
myMap <- get_map(location=myLocation,
source="stamen", maptype="watercolor", crop=FALSE)
```

#Same camera location plot but using ggmap
```{r}
ggmap(myMap)+
geom_point(aes(x = Longitude, y = Latitude), data = cameras,
 alpha = .5, color="darkred", size = 0.5) +
ggtitle("Traffic Cameras in DC")
```

#Plotting all of the crashes in DC from 2016-2019
```{r}
ggmap(myMap)+
geom_point(aes(x = LONGITUDE, y = LATITUDE), data = crashes,
 alpha = .25, color="darkred", size = 0.5) +
ggtitle("Traffic Crashes in DC")
```

#Sampling Crashes
```{r}
crashes1<-crashes[sample(nrow(crashes),1000),]
```

#Plotting the sample
```{r}
ggmap(myMap)+
geom_point(aes(x = LONGITUDE, y = LATITUDE), data = crashes1,
 alpha = .5, color="darkred", size = 0.5) +
ggtitle("Traffic Crashes in DC")
```
#Creating Clusters for Entire Crash dataset
```{r}
set.seed(20)
clusters <- kmeans(crashes[,1:2], 7)
crashes$ward_c <- as.factor(clusters$cluster)
```

#Inspect the Clusters
```{r}
str(clusters)
```

#Plot the clusters
```{r}
ggmap(myMap) + geom_point(aes(x = LONGITUDE[], y = LATITUDE[], colour = as.factor(ward_c)),data = crashes) +
  ggtitle("Crash Clusters using KMeans")
```

#Creating Clusters for Crash Sample
```{r}
set.seed(20)
clusters <- kmeans(crashes1[,1:2], 7)
crashes1$ward_c <- as.factor(clusters$cluster)
```

#Plot the clusters for Crash Sample
```{r}
ggmap(myMap) + geom_point(aes(x = LONGITUDE[], y = LATITUDE[], colour = as.factor(ward_c)),data = crashes1) +
  ggtitle("Crash Clusters using KMeans for Sample")
```
- Clusters are much easier to point out here when taken from the sample.

#Create a dummy variable for the sample to join with violations
```{r}
crashes1$TYPE<-0
crashes1$TYPE<-as.factor(crashes1$TYPE)
```

