---
title: "Untitled"
output: html_document
---
#Loading relevant libraries
```{r}
library(rgdal)
library(raster)
library(ggmap)
library(lubridate)
library(sf)
library(ggplot2)
library(maps)
library(mapdata)
library(sp)
```

####Reading 2016
```{r}
#Reading in Violations for 2016
violation.jan.2016<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_January_2016.csv")
violation.feb.2016<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_February_2016.csv")
violation.mar.2016<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_March_2016.csv")
violation.april.2016<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_April_2016.csv")
violation.may.2016<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_May_2016.csv")
violation.june.2016<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_June_2016.csv")
violation.july.2016<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_July_2016.csv")
violation.aug.2016<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_August_2016.csv")
violation.sep.2016<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_September_2016.csv")
violation.oct.2016<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_October_2016.csv")
violation.nov.2016<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_November_2016.csv")
violation.dec.2016<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_December_2016.csv")
```
####Binding 2016
```{r}
#Binding 2016 Violations to create one vertical table
violations2016a<-rbind(violation.jan.2016, violation.feb.2016)
violations2016b<-rbind(violations2016a, violation.mar.2016)
violations2016c<-rbind(violations2016b, violation.april.2016)
violations2016d<-rbind(violations2016c, violation.may.2016)
violations2016e<-rbind(violations2016d, violation.june.2016)
violations2016f<-rbind(violations2016e, violation.july.2016)
violations2016g<-rbind(violations2016f, violation.aug.2016)
violations2016h<-rbind(violations2016g, violation.sep.2016)
violations2016i<-rbind(violations2016h, violation.oct.2016)
violations2016j<-rbind(violations2016i, violation.nov.2016)
violations2016k<-rbind(violations2016j, violation.dec.2016)
#Rename Violations 2016 data
violations2016<-violations2016k
```

####Reading 2017
```{r}
#Reading in Violations for 2017
violation.jan.2017<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_January_2017.csv")
violation.feb.2017<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_February_2017.csv")
violation.mar.2017<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_March_2017.csv")
violation.april.2017<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_April_2017.csv")
violation.may.2017<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_May_2017.csv")
violation.june.2017<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_June_2017.csv")
violation.july.2017<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_July_2017.csv")
violation.aug.2017<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_August_2017.csv")
violation.sep.2017<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_September_2017.csv")
violation.oct.2017<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_October_2017.csv")
violation.nov.2017<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_November_2017.csv")
violation.dec.2017<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_December_2017.csv")
```
####Binding 2017
```{r}
#Binding 2017 Violations to create one vertical table
violations2017a<-rbind(violation.jan.2017, violation.feb.2017)
violations2017b<-rbind(violations2017a, violation.mar.2017)
violations2017c<-rbind(violations2017b, violation.april.2017)
violations2017d<-rbind(violations2017c, violation.may.2017)
violations2017e<-rbind(violations2017d, violation.june.2017)
violations2017f<-rbind(violations2017e, violation.july.2017)
violations2017g<-rbind(violations2017f, violation.aug.2017)
violations2017h<-rbind(violations2017g, violation.sep.2017)
violations2017i<-rbind(violations2017h, violation.oct.2017)
violations2017j<-rbind(violations2017i, violation.nov.2017)
violations2017k<-rbind(violations2017j, violation.dec.2017)
#Rename Violations 2017 data
violations2017<-violations2017k
```

####Reading 2018
```{r}
#Reading in Violations for 2018
violation.jan.2018<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_January_2018.csv")
violation.feb.2018<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_February_2018.csv")
violation.mar.2018<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_March_2018.csv")
violation.april.2018<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_April_2018.csv")
violation.may.2018<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_May_2018.csv")
violation.june.2018<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_June_2018.csv")
violation.july.2018<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_July_2018.csv")
violation.aug.2018<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_August_2018.csv")
violation.sep.2018<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_September_2018.csv")
violation.oct.2018<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_October_2018.csv")
#
violation.nov.2018<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_November_2018.csv")
violation.dec.2018<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_December_2018.csv")
```
####Binding 2018
```{r}
#Quick Preclean for unneccessary column
violation.nov.2018$ROW_<-NULL
#Binding 2018 Violations to create one vertical table
violations2018a<-rbind(violation.jan.2018, violation.feb.2018)
violations2018b<-rbind(violations2018a, violation.mar.2018)
violations2018c<-rbind(violations2018b, violation.april.2018)
violations2018d<-rbind(violations2018c, violation.may.2018)
violations2018e<-rbind(violations2018d, violation.june.2018)
violations2018f<-rbind(violations2018e, violation.july.2018)
violations2018g<-rbind(violations2018f, violation.aug.2018)
violations2018h<-rbind(violations2018g, violation.sep.2018)
violations2018i<-rbind(violations2018h, violation.oct.2018)
violations2018j<-rbind(violations2018i, violation.nov.2018)
violations2018k<-rbind(violations2018j, violation.dec.2018)
#Rename Violations 2018 data
violations2018<-violations2018k
```


####Reading 2019
```{r}
#Reading in Violations for 2019
violation.jan.2019<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_January_2019.csv")
violation.feb.2019<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_February_2019.csv")
violation.mar.2019<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_March_2019.csv")
violation.april.2019<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_April_2019.csv")
violation.may.2019<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_May_2019.csv")
violation.june.2019<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_June_2019.csv")
violation.july.2019<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_July_2019.csv")
violation.aug.2019<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_August_2019.csv")
violation.sep.2019<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_September_2019.csv")
violation.oct.2019<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_October_2019.csv")
violation.nov.2019<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_November_2019.csv")
violation.dec.2019<-read.csv("/Users/BeauSwallow/Downloads/Moving_Violations_Issued_in_December_2019.csv")
```
####Precleaning 2019 for binding, this variable is only in July 2019 onward and isn't useful
```{r}
violation.july.2019$VIOLATION_TYPE_DESC<-NULL
violation.aug.2019$VIOLATION_TYPE_DESC<-NULL
violation.sep.2019$VIOLATION_TYPE_DESC<-NULL
violation.oct.2019$VIOLATION_TYPE_DESC<-NULL
violation.nov.2019$VIOLATION_TYPE_DESC<-NULL
violation.dec.2019$VIOLATION_TYPE_DESC<-NULL
```

####Binding 2019
```{r}
#Binding 2019 Violations to create one vertical table
violations2019a<-rbind(violation.jan.2019, violation.feb.2019)
violations2019b<-rbind(violations2019a, violation.mar.2019)
violations2019c<-rbind(violations2019b, violation.april.2019)
violations2019d<-rbind(violations2019c, violation.may.2019)
violations2019e<-rbind(violations2019d, violation.june.2019)
violations2019f<-rbind(violations2019e, violation.july.2019)
violations2019g<-rbind(violations2019f, violation.aug.2019)
violations2019h<-rbind(violations2019g, violation.sep.2019)
violations2019i<-rbind(violations2019h, violation.oct.2019)
violations2019j<-rbind(violations2019i, violation.nov.2019)
violations2019k<-rbind(violations2019j, violation.dec.2019)
#Rename Violations 2019 data
violations2019<-violations2019k
```

#Cleaning the dates for all 4
```{r}
#Making Date Column cleaner
violations2016$ISSUE_DATE<-as.Date(violations2016$ISSUE_DATE)
violations2017$ISSUE_DATE<-as.Date(violations2017$ISSUE_DATE)
#violations2018$ISSUE_DATE<-as.Date(violations2018$ISSUE_DATE)
violations2019$ISSUE_DATE<-as.Date(violations2019$ISSUE_DATE)
#Removing unneccessary columns for 2016
violations2016$ISSUING_AGENCY_CODE<-NULL
violations2016$ISSUING_AGENCY_NAME<-NULL
violations2016$ISSUING_AGENCY_SHORT<-NULL
violations2016$DISPOSITION_CODE<-NULL
violations2016$DISPOSITION_TYPE<-NULL
violations2016$DISPOSITION_DATE<-NULL
violations2016$FINE_AMOUNT<-NULL
violations2016$TOTAL_PAID<-NULL
violations2016$PENALTY_1<-NULL
violations2016$PENALTY_2<-NULL
violations2016$PENALTY_3<-NULL
violations2016$PENALTY_4<-NULL
violations2016$PENALTY_5<-NULL
violations2016$RP_MULT_OWNER_NO<-NULL
violations2016$GIS_LAST_MOD_DTTM<-NULL
#Removing unneccessary columns for 2017
violations2017$ISSUING_AGENCY_CODE<-NULL
violations2017$ISSUING_AGENCY_NAME<-NULL
violations2017$ISSUING_AGENCY_SHORT<-NULL
violations2017$DISPOSITION_CODE<-NULL
violations2017$DISPOSITION_TYPE<-NULL
violations2017$DISPOSITION_DATE<-NULL
violations2017$FINE_AMOUNT<-NULL
violations2017$TOTAL_PAID<-NULL
violations2017$PENALTY_1<-NULL
violations2017$PENALTY_2<-NULL
violations2017$PENALTY_3<-NULL
violations2017$PENALTY_4<-NULL
violations2017$PENALTY_5<-NULL
violations2017$RP_MULT_OWNER_NO<-NULL
violations2017$GIS_LAST_MOD_DTTM<-NULL
#Removing unneccessary columns for 2018
#violations2018$ISSUING_AGENCY_CODE<-NULL
#violations2018$ISSUING_AGENCY_NAME<-NULL
#violations2018$ISSUING_AGENCY_SHORT<-NULL
#violations2018$DISPOSITION_CODE<-NULL
#violations2018$DISPOSITION_TYPE<-NULL
#violations2018$DISPOSITION_DATE<-NULL
#violations2018$FINE_AMOUNT<-NULL
#violations2018$TOTAL_PAID<-NULL
#violations20187$PENALTY_1<-NULL
#violations2018$PENALTY_2<-NULL
#violations2018$PENALTY_3<-NULL
#violations2018$PENALTY_4<-NULL
#violations2018$PENALTY_5<-NULL
#violations2018$RP_MULT_OWNER_NO<-NULL
#violations2018$GIS_LAST_MOD_DTTM<-NULL
#Removing unneccessary columns for2019
violations2019$ISSUING_AGENCY_CODE<-NULL
violations2019$ISSUING_AGENCY_NAME<-NULL
violations2019$ISSUING_AGENCY_SHORT<-NULL
violations2019$DISPOSITION_CODE<-NULL
violations2019$DISPOSITION_TYPE<-NULL
violations2019$DISPOSITION_DATE<-NULL
violations2019$FINE_AMOUNT<-NULL
violations2019$TOTAL_PAID<-NULL
violations2019$PENALTY_1<-NULL
violations2019$PENALTY_2<-NULL
violations2019$PENALTY_3<-NULL
violations2019$PENALTY_4<-NULL
violations2019$PENALTY_5<-NULL
violations2019$RP_MULT_OWNER_NO<-NULL
violations2019$GIS_LAST_MOD_DTTM<-NULL
```


#Reordering columns so Latitude and Longitude are columns 1 and 2
```{r}
violations20162 <- violations2016[, c(13, 14, 1, 2, 3, 4, 5, 6 ,7, 8, 9, 10, 11, 12, 13, 15)]
violations2016<-violations20162
violations20172 <- violations2017[, c(13, 14, 1, 2, 3, 4, 5, 6 ,7, 8, 9, 10, 11, 12, 13, 15)]
violations2017<-violations20172
#violations20182 <- violations2018[, c(13, 14, 1, 2, 3, 4, 5, 6 ,7, 8, 9, 10, 11, 12, 13, 15)]
#violations2018<-violations20182
violations20192 <- violations2019[, c(11, 10, 1, 2, 3, 4, 5, 6 ,7, 8, 9, 12, 13, 14, 15)]
violations2019<-violations20192
```

#head(violations20182)

#Eliminating NA/NAN/Inf values for kmeans() analysis
```{r}
violations2016<-na.omit(violations2016)
violations2017<-na.omit(violations2017)
#violations2018<-na.omit(violations2018)
violations2019<-na.omit(violations2019)
```


#Summary Statistics
- Changed all functions to head() so it would knit correctly.
```{r}
head(violations2016)
#1,154,050 violations for 2016 with NA
#1,033,269 violations for 2016 omitting NA
```


```{r}
head(violations2017)
#1,337,742 violations for 2017
#1,201,375 violations for 2017 omitting NA
```

```{r}
#violations2018
#1,XXX,XXX violations for 2018
#1,XXX,XXX violations for 2018 omitting NA
```

```{r}
head(violations2019)
#1,287,967 violations for 2019
#919,606  violations for 2019 omitting NA
```

#Merging the four years of datasets
```{r}
#violations_a<-rbind(violations2016, violations2017)
#violations_b<-rbind(violations_a, violations2018)
#violations_c<-rbind(violations_b, violations2019)
#Renaming
#violations<-violations_c
```

#Sample the four years of violations
```{r}
violation_sample_2016<-violations2016[sample(nrow(violations2016),1000),]
violation_sample_2017<-violations2017[sample(nrow(violations2016),1000),]
#violation_sample_2018<-violations2018[sample(nrow(violations2016),1000),]
violation_sample_2019<-violations2019[sample(nrow(violations2016),1000),]
```

#DC map using ggmap library
```{r}
myLocation <- c(-77.1198, 38.79164, -76.90915, 38.99597)
myMap <- get_map(location=myLocation,
source="stamen", maptype="watercolor", crop=FALSE)
```
#Plotting 2016 Traffic Violations
```{r}
ggmap(myMap)+
geom_point(aes(x = LONGITUDE, y = LATITUDE), data = violations2016,
 alpha = .25, color="darkred", size = 1) +
ggtitle("Traffic Violations in DC 2016")
```

#Plotting Sample of 2016 Traffic Violations
```{r}
ggmap(myMap)+
geom_point(aes(x = LONGITUDE, y = LATITUDE), data = violation_sample_2016,
 alpha = .5, color="darkred", size = 1) +
ggtitle("Sample of Traffic Violations in DC 2016")
```

#Creating Clusters for the 2016 Violations Sample dataset
```{r}
set.seed(20)
clusters <- kmeans(violation_sample_2016[,1:2], 7)
violation_sample_2016$ward_c <- as.factor(clusters$cluster)
```

#Plotting the Clusters for the  2016 Violations Sample dataset
```{r}
ggmap(myMap) + geom_point(aes(x = LONGITUDE[], y = LATITUDE[], colour = as.factor(ward_c)),data = violation_sample_2016) +
  ggtitle("Clusters for Sample of 2016 Violations using KMeans")
```


